Lab2

### 2021-08-04 src/data/DataCleaning.Rmd
limited model dataframe completed

```{r import libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
# source(local=TRUE, file="../util.R")
```

```{r read in data, echo=FALSE}
d_covid <- read.csv(url('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'))
d_openings <- read.csv(file = './../../data/raw/COVID-19 US state policy database (CUSP) - Closures & Reopening.csv')
d_trips <- read.csv(file = './../../data/raw/Trips_by_Distance.csv')
```

@@

# dates under investigation
start_date = as.Date("2021-06-01")
end_date = as.Date("2021-06-30")

# start_date = as.Date("2021-06-01")
# end_date = as.Date("2021-06-30")
```

@@

# cleaning
d_trips <- d_trips %>%
  # clean county name format
  mutate(
    County.Name = gsub("([A-Za-z]+).*", "\\1", County.Name),
    total_pop = select(.,Population.Staying.at.Home:Population.Not.Staying.at.Home) %>% rowSums(na.rm = TRUE)
    ) %>%
  group_by(County.FIPS) %>%
  mutate(
    # define long trips
    long_trip = sum(c_across(Number.of.Trips.100.250:Number.of.Trips...500)),
    long_trip_pc = long_trip/total_pop,
    # define short trips
    short_trip = sum(c_across(Number.of.Trips.1.3:Number.of.Trips.50.100)),
    short_trip_pc = short_trip/total_pop
  ) %>%
  rename(
    state = State.Postal.Code,
    county = County.Name,
    fips = County.FIPS
  ) %>%
  select(
    county, fips, state, total_pop, long_trip, long_trip_pc, short_trip, short_trip_pc
  ) %>%
  distinct(fips, .keep_all = TRUE)
```

@@

# Clean other variables
```{r read in variable data, echo=FALSE}
density <- read.delim("./../../data/raw/2019_Gaz_counties_national.txt", header=TRUE, sep="\t")
dp03_econ <- read.csv(file = './../../data/raw/ACSDP5Y2019.DP03_data_with_overlays_2021-07-26T143523.csv')
dp05_age <- read.csv(file = './../../data/raw/ACSDP1Y2019.DP05_data_with_overlays_2021-07-21T215013.csv')
s1901_income <- read.csv(file = './../../data/raw/ACSST1Y2019.S1901_data_with_overlays_2021-07-21T222906.csv') 
urb_rur_h2 <- read.csv(file = './../../data/raw/DECENNIALSF12010.H2_data_with_overlays_2021-06-08T173848.csv')

unemp_claims <- readxl::read_excel("./../../data/raw/weekly-claims-by-county-twc.xlsx")

@@

# merge into main dataframe
d <- merge(x = econ, y = d, by = c("fips"))
```
```{r clean twc unemployment data, echo=FALSE}

@@

>>>>

```{r import libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

# limited model data cleaning

```{r read in data, echo=FALSE}
d_covid <- read.csv(url('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'))
d_trips <- read.csv(file = './../../data/raw/Trips_by_Distance.csv')
urb_rur_h2 <- read.csv(file = './../../data/raw/DECENNIALSF12010.H2_data_with_overlays_2021-06-08T173848.csv') 
density <- read.csv(file='./../../data/raw/TAC_population.csv')
income <- read.csv('./../../data/raw/TAC_income.csv')
```

@@

# dates under investigation
start_date = as.Date("2021-06-01")
end_date = as.Date("2021-06-30")
```

@@

# cleaning/county filter
d_trips <- d_trips %>%
  # clean county name format
  mutate(
    County.Name = gsub("([A-Za-z]+).*", "\\1", County.Name),
    total_pop = select(.,Population.Staying.at.Home:Population.Not.Staying.at.Home) %>% rowSums(na.rm = TRUE)
    ) %>%
  group_by(County.FIPS) %>%
  mutate(
    # define long trips
    long_trip = sum(c_across(Number.of.Trips.100.250:Number.of.Trips...500)),
    # define short trips
    # short_trip = sum(c_across(Number.of.Trips.1.3:Number.of.Trips.50.100)),
  ) %>%
  rename(
    state = State.Postal.Code,
    county = County.Name,
    fips = County.FIPS
  ) %>%
  select(
    county, fips, state, total_pop, long_trip
  ) %>%
  distinct(fips, .keep_all = TRUE)
```

```{r clean urban/rural data}
rural_ind <- urb_rur_h2[-c(1),] %>%
  rename(fips = GEO_ID,
         total_pop = H002001,
         urbanized = H002003,
         rural = H002005) %>%
  select(fips, total_pop, urbanized, rural) %>%
  mutate(
    fips = substring(fips, 10),
    total_pop = as.double(total_pop),
    urbanized = as.double(urbanized),
    rural = as.double(rural),
    is_rural = ifelse((total_pop < 50000 & urbanized == 0), 1, 0)
  ) %>% 
  select(fips, is_rural) %>% drop_na()
```

```{r clean income data}
income_f <- income %>%
  rename(
    county = County,
    median_income = Median.Household.Income,
    povery_rate = X..of.Population.in.Poverty
  ) %>%
  select(
    county, median_income, povery_rate
  ) %>%
  mutate(
    median_income = as.double(gsub('[$,]', '', median_income))
  )

```{r clean density data}
density_f <- density %>%
  rename(
    county = County,
    total_pop = County.Population,
    density = X.Population.Density.Per.Sq.Mile
  ) %>%
  select(
    county, total_pop, density
  ) %>%
  mutate(
    total_pop = as.double(gsub('[,]', '', total_pop)),
    density = as.double(gsub('[,]', '', density))
  )

```{r merge limited model dataframe}
d <- merge(x = d_trips, y = d_covid, by = c("fips", "county", "state")) 
d <- merge(x = rural_ind, y = d, by = c("fips"))
d <- merge(x = income_f, y = d, by = c("county"))
d <- merge(x = density_f, y = d, by = c("county"))
d <- d %>% drop_na()

#rm(d)
```

@@

# Model 2 data cleaning

# Model 3 data cleaning

# Clean other variables
```{r read in variable data, echo=FALSE}
density <- read.delim("./../../data/raw/2019_Gaz_counties_national.txt", header=TRUE, sep="\t")
dp03_econ <- read.csv(file = './../../data/raw/ACSDP5Y2019.DP03_data_with_overlays_2021-07-26T143523.csv')
dp05_age <- read.csv(file = './../../data/raw/ACSDP1Y2019.DP05_data_with_overlays_2021-07-21T215013.csv')
s1901_income <- read.csv(file = './../../data/raw/ACSST1Y2019.S1901_data_with_overlays_2021-07-21T222906.csv')

unemp_claims <- readxl::read_excel("./../../data/raw/weekly-claims-by-county-twc.xlsx")
ur <- readxl::read_excel("./../../data/raw/TX_Rural_County.xlsx")
```

@@

# merge into main dataframe
d <- merge(x = econ, y = d, by = c("fips"))
```

```{r clean twc unemployment data, echo=FALSE}

@@

```{r rural indicator, echo=FALSE}
d$rural_ind <- ifelse(d$county %in% ur$`Rural County`, 1, 0)

summary(d$rural_ind)
